Demo: Counterpoint by Construction

### 2019-05-25 doc/farm/farm-abstract.tex
added some rules of first species counterpoint

In this demonstration of work in progress, we present Music Tools
\citep{MusicTools}, a library of small tools that can be combined
functionally to help analyze and synthesize music.
To allow simple and natural encoding of rules, we built the library
in Agda \citep{norellphd}, which is a functional language with full
dependent types.
As an application of the library, we demonstrate an implementation of 
species counterpoint, based on the rules given by \citet{fux1965study}.
Thanks to Agda's rich type system, we can ensure by construction that
well-typed counterpoint satisfies all the required rules.
We show how the type-based approach both aids human composition and
allows for computer-generated creation of correct counterpoint.
We also contrast our work to a recent study on generating
natural-sounding counterpoint by machine learning
\citep{CounterpointByConvolution}, which does not give us correctness
guarantees.
Finally, we discuss further applications of our library, including
handling of functional harmony.

>>>>

In this demonstration of work in progress, we present Music Tools
\citep{MusicTools}, a library of small tools that can be combined
functionally to help analyze and synthesize music.
To allow simple and natural encoding of rules, we built the library
in Agda \citep{norellphd}, which is a functional language with full
dependent types.
As an application of the library, we demonstrate an implementation of 
species counterpoint, based on the rules given by \citet{fux1965study}.
To give a rough idea of what the rules look like, here are some of
the requirements for composing two-part, first-species counterpoint:

\begin{itemize}
\item The interval between the two voices should not be a perfect fourth
\item The two voices should not move to a perfect fifth or octave 
by similar motion
\item The counterpoint should end with a proper cadence
\end{itemize}

\noindent Thanks to Agda's rich type system, we can fully express 
these rules using types, and thus ensure by construction that
well-typed counterpoint satisfies all the required rules.

At the FARM workshop, we intend to give a gentle introduction to
counterpoint, and describe our Agda implementation, showing how the
type-based approach both aids human composition and allows for 
computer-generated creation of correct counterpoint.
We then contrast our work to a recent study on generating
natural-sounding counterpoint by machine learning
\citep{CounterpointByConvolution}, which does not give us correctness
guarantees.
Finally, we discuss further applications of our library, including
handling of functional harmony.
